\propertitle

### 2019-07-07 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-25T17:40:21+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The maximum-entropy method has been applied in the neurosciences to
questions of sufficient statistics -- for example, whether we approximately
reach the same state of uncertainty about network activity by considering
only its network-averaged pairwise correlations and by considering also
third-order correlations. The method is typically used as follows. We
imagine to be ignorant of the measured frequency distribution for the
network activity, and to know only the measured statistics of interest; two
maximum-entropy distributions are then built, constrained under the two
sets of moments under comparison. Finally, the resulting distributions are
compared with each other and with the measured frequency distribution.

The method introduced in the present work reveals an issue in the recipe
above. The issue doesn't come from the method itself but from a simple,
general fact of probability theory: if a probability distribution has some
sufficient statistics \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000},
then its marginals \emph{cannot} have the same sufficient statistics, and
vice versa (except for trivial cases, such as uniform distributions). This
also holds among marginals of different order (to $n$ neurons or to $n+1$
neurons, for example). This impossibility is known in statistical
mechanics: marginals of Gibbs states aren't Gibbs states
\citep{maesetal1999}. Mathematically it corresponds to the general
impossibility of solving a system of $n$ equations in $m$ unknowns when
$n>m$ \citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.

In the neuroscientific application above, this means for example that if
pairwise correlations are sufficient statistics for a sample of a network,
then they \emph{cannot} be sufficient for the full network or for a sample
with more or fewer neurons; and vice versa, if we assume that they are
sufficient statistics for the full network, then we must expect them
\emph{not} to be sufficient for any recorded sample thereof.

It seems to us that such questions about sufficient statistics are
implicitly asked for specific brain areas, not for casually recorded
samples thereof. Thus, maximum-entropy methods applied at the sample level
paradoxically give an answer opposite to the one we might think they're
giving!

@@

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-07-07T14:24:05+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The maximum-entropy method has often been applied in the neurosciences to
questions of sufficient statistics. For example: does knowledge about
third-order correlations change our state of uncertainty about network
activity in an appreciable way, compared to knowledge about second-order
correlations only? The method is used as follows to answer this question.
We imagine to be ignorant of the measured frequency distribution for the
network activity, and to know only up to second-order measured statistics,
or up to third-order. Two maximum-entropy distributions are then built,
constrained under the two sets of moments under comparison. Finally, the
resulting distributions are compared with each other and with the measured
frequency distribution. This recipe generalizes to any subsets of measured
statistics, of course.

The method introduced in the present work reveals an issue in this recipe;
the issue doesn't come from the method itself but from a simple, general
fact of probability theory. If a probability distribution has some
sufficient statistics \citep[\sect~4.5]{bernardoetal1994}{fortinietal2000},
then its marginals \emph{cannot} have the same sufficient statistics, and
vice versa, except for trivial cases such as uniform distributions. This in
general also holds between different marginals: if one has some sufficient
statistics, another won't have it. This impossibility is known in
statistical mechanics: marginals of Gibbs states aren't Gibbs states
\citep{maesetal1999}. Mathematically it corresponds to the general
impossibility of solving a system of $n$ equations in $m$ unknowns when
$n>m$ \citep[\sect~3.1]{portamanaetal2015}[\sect~3]{portamanaetal2018b}.

In the neuroscientific application above, this means for example that if
pairwise correlations are sufficient statistics for a particular sample of
a network, then they \emph{cannot} be sufficient for the full networ, or
for a different sample; and vice versa, if we assume that they are
sufficient statistics for the full network, then we must expect them
\emph{not} to be sufficient for any recorded sample thereof. But questions
about sufficient statistics are implicitly meant to apply to a whole brain
area, not just to some casually recorded sample. Thus, maximum-entropy
methods applied at the sample level paradoxically give an answer opposite
to the one we might think they're giving!

The comparisons of different sufficient statistics are only approximate,
however: we ask only whether our uncertainties conditional on either are
\emph{appreciably} different in some specified sense. It's therefore worth
checking whether the different results we get from a sufficient statistics
for the sample and for the full-network are still very similar when we
consider an altogether different statistics.

@@

%plot_mc_6mom_unif_afss3tr_N10000_t400_L500_T5e+05_s30000_a0_i3
